Exploratory Data Visualizations (Fig. 1)

### 2021-02-16 analysis/paper/EDA_figure_1.Rmd
added a survivorship curve for the 1998 cohort of plants

```{r}
demog_full <- read_rds(here("analysis", "data", "derived_data", "ha_spei_combined.rds"))
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
```

@@

```{r}
demog_full <-
  demog_full %>% 
  filter(ht < 200 | is.na(ht), ht_prev < 200 | is.na(ht_prev)) %>%
  filter(
    # having no aboveground biomass is possible, but categorically different than just being small. Exclude plants with 0 shoots or 0 height
    shts_prev > 0,
    ht_prev > 0,
    #for survival data, must include plants with NA for shts and height (i.e. dead plants)
    shts > 0 | is.na(shts), 
    ht > 0 | is.na(ht)
  ) %>%
  mutate(size = shts*ht, size_prev = shts_prev * ht_prev, log_size = log(size), log_size_prev = log(size_prev))
```

@@

@@

```{r}
size/survival/flowering/spei +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(plot.margin = margin(2,1,1,1)) &
  guides(col = guide_legend(title = "Habitat"))
ggsave(here("analysis", "figures", "eda.png"), height = 7, width = 10)
```

>>>>

```{r}
demog <- read_rds(here("analysis", "data", "derived_data", "ha_spei_combined.rds"))
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
```

@@

```{r}
demog_full <-
  demog %>% 
  filter(ht < 200 | is.na(ht), ht_prev < 200 | is.na(ht_prev)) %>%
  filter(
    # having no aboveground biomass is possible, but categorically different than just being small. Exclude plants with 0 shoots or 0 height
    shts_prev > 0,
    ht_prev > 0,
    #for survival data, must include plants with NA for shts and height (i.e. dead plants)
    shts > 0 | is.na(shts), 
    ht > 0 | is.na(ht)
  ) %>%
  mutate(size = shts*ht, size_prev = shts_prev * ht_prev, log_size = log(size), log_size_prev = log(size_prev))
```

@@

### Survivorship curve

Emilio suggested an alternative to the timeseries plot---a survivorship curve for the 1998 cohort of plants.

```{r paged.print=FALSE}
cohort <-demog %>% filter(year == 1998) %>% pull(ha_id_number) %>% unique()

surv_curve_df <-
  demog %>% 
  filter(ha_id_number %in% cohort) %>% 
  filter(surv == 1, habitat != "10-ha") %>% 
  group_by(year, habitat) %>% 
  summarize(n = n()) %>%
  group_by(habitat) %>% 
  mutate(p_surv = n/first(n),
         date = ymd(paste0(year, "-02-01")),
         yearmonth = yearmonth(date))

surv_curve <-
  ggplot(surv_curve_df, aes(x = yearmonth, y = p_surv, color = habitat)) +
  geom_step() +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(survived)") +
  theme_bw() +
  theme(
    # axis.text.x = element_blank(),
    # axis.title.x = element_blank(),
    # axis.ticks.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
surv_curve
```

@@

# Combine

```{r}
size/
  # survival/
  surv_curve/
  flowering/
  spei +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(plot.margin = margin(2,1,1,1)) &
  guides(col = guide_legend(title = "Habitat"))
ggsave(here("analysis", "figures", "eda.png"), height = 7, width = 10)
```
